---
title: "BioMove swallow-parasite manuscript (4)" ## name of your project and analysis step
description: "continuous-time-movement models (ctmm)"
author:
    - name: "Marius Grabow"
      email: "grabow@izw-berlin.de" 
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
      orcid_id: 0000-0003-0702-9642
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: true  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

# Setup
Load required packages

```{r packages, message=FALSE}
library("here")
library("tidyverse")
library("ctmm")
library("sf")
library("foreach")
```

# Background:

Here, we fit time continuous movement models to ATLAS movement data of barn swallows and house martins. We fit separate models, (A) first for each individual that was tagged, (B) second, we fit models for each individual and day to later account for the variation between days.

# CTMM workflow
## 1 Prepare data


- drop animals without known infection status (i.e. animals that were not tested)
- rename columns to movebank conventions
- transform to telemetry object
- select important covariates to keep

```{r}
# create prep function
prep<-function(df,TIME){
  
  
  df%>%
  sf::st_drop_geometry() %>%
  # filter out all animals that have not been blood-sampled
  filter(!is.na(blood_infection)) %>%
  #rename
  mutate(
    sensor = "atlas",
    tag_id = as.character(tag_id),
    TAG = tag_id) %>%
  #doesn't like tibbles
  as.data.frame() %>%
  #rename
  rename(
    individual.local.identifier = tag_id,
    timestamp = TIME,
    location.long = long,
    location.lat = lat,
    COV.x.y = COVXY,
    VAR.x = VARX,
    VAR.y = VARY
  ) %>%
  #telemetry object
  as.telemetry(
    projection = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs",
    keep = c(
      "TAG",
      "species",
      "sex",
      "COV.x.y",
      "VAR.x",
      "VAR.y",
      "blood_infection",
      "sum_infection",
      "infection_type",
      "plasmodium",
      "leucocytozoon",
      "haemoproteus"
    )
  )%>%
    keep( ~ nrow(.x) > 500)
}

# Function to realign timestamps
realign_timestamps <- function(df, interval) {
  df %>%
    mutate(start_time = first(TIME), # Dynamically determine the start time for each group
           interval_count = round((TIME - start_time) / interval),
           adjusted_timestamp = start_time + interval_count * interval)
}
```

```{r}
##apply function
swallows_tag <- readRDS(here("output",
                             "data-proc",
                             "swallows_day_all.RDS")) %>%
  dplyr::filter(!is.na(blood_infection)) %>%
  group_by(tag_id) %>%
  realign_timestamps(interval = 8) %>%
  mutate(TIME = adjusted_timestamp) %>%
  as.data.frame() %>%
  distinct(TIME, .keep_all = TRUE) %>%
  ungroup() %>%
  prep(TIME)


names(swallows_tag) -> keep

uere(swallows_tag) <- 1

### save true locations
true_locations <- swallows_tag %>%
  map_df( ~ enframe(.x$t, name = NULL, value = "t"), .id = "tag_id") %>%
  mutate(location = "real")

true <- split(true_locations$t, true_locations$tag_id)

```


## 2 Fit CTMM for each individual

Parallelize (Don't run this code on a local machine)


```{r tag ctmm fit, eval=FALSE, include=TRUE}

# set-up parallel sessions
cl <- parallel::makeCluster(20)
doParallel::registerDoParallel(cl)

# create function to fit data & select best movement model
model_fit <- function(i)
{
  GUESS <-
    ctmm.guess(swallows_tag[[i]],
               CTMM = ctmm(error = TRUE),
               interactive = FALSE)
  ctmm.select(swallows_tag[[i]], GUESS)
}

#iterate over list
ctmm_fit_tag_id <-
  foreach(
    i = 1:length(swallows_tag),
    .packages = 'ctmm',
    .errorhandling = 'remove'
  ) %dopar% {
    model_fit(i)
  }

#same names
names(ctmm_fit_tag_id) <- names(swallows_tag)

## 4 Plot Variograms

```{r tag variograms, cache=TRUE}
level <- c(0.5, 0.95)
# PLot all variogramms

for (i in 1:length(swallows_tag)) {
  VARIOGRAM <- variogram(swallows_tag[[i]])
  
  PLOT <-
    plot(
      VARIOGRAM,
      CTMM = ctmm_fit_tag[[i]],
      fraction = 0.7,
      level = level,
      col.CTMM = "blue"
    )
  
  
  print(PLOT)
  
  title(
    main = paste(
      names(swallows_tag[i]),
      "species = ",
      swallows_tag[[i]]$species[1],
      "sex = ",
      swallows_tag[[i]]$sex[1],
      "blood infection = ",
      swallows_tag[[i]]$blood_infection[1]
    )
  )
}
```

## 5 Calculate AKDEs by TAG

```{r calculate tag akde, include=TRUE, eval=FALSE}

# set-up parallel session
cl <- parallel::makeCluster(50)
doParallel::registerDoParallel(cl)

# function for calculating akde for each individual (using the ctmm model)
akde_fit_weights <- function(i)
{
  AKDE <- akde(swallows_tag[[i]],
               ctmm_fit_tag[[i]],
               weights = F)
}

#iterate over list
AKDES_tag_weights <- foreach(i=1: length(swallows_tag),.packages='ctmm',.errorhandling = 'remove') %dopar% { akde_fit_weights(i) }

names(AKDES_tag_weights)<-names(swallows_tag)

same_grid_akde<-akde(swallows_tag,ctmm_fit_tag_id)

```

## 6 Calculate akde sizes as data.frame

```{r}

AKDE_size<-NULL

for(i in 1:length(akdes_tag_2024)){
  TAG <- names(akdes_tag_2024[i])
  min_AREA <- summary(akdes_tag_2024[[i]],units=F)$CI[1]
  mean_AREA <- summary(akdes_tag_2024[[i]],units=F)$CI[2]
  max_AREA <- summary(akdes_tag_2024[[i]],units=F)$CI[3]
  blood_infection <- swallows_tag[[i]]$blood_infection[1]
  species <- swallows_tag[[i]]$species[1]
  
  AKDE_size = rbind(AKDE_size, data.frame(TAG,min_AREA,mean_AREA,max_AREA,blood_infection,species))
  
}

```


## 7.1 Meta Analysis Barn swallow & infection

Here, we perform a meta analysis based on chi-square statistics (see Fleming et al. 2022)

```{r Barn swallow & infection, cache=TRUE}

healthy_barn <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "RS" & blood_infection == "n")%>%
  pull(index)

infected_barn <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "RS" & blood_infection == "y")%>%
  pull(index)

# meta analysis
# akde 95%
(meta(list(BS_NEG=
            akdes_tag_2024[healthy_barn],
          BS_POS=
            akdes_tag_2024[infected_barn]),
      units=TRUE,
     plot = TRUE,
     verbose = TRUE,
     level.UD = 0.95,
     level = 0.95,debias = F,
     main="homerange 95% AKDE isopleth")->RS_meta_export_tag)

# akde 50%
meta(list(BS_NEG=
            akdes_tag_2024[healthy_barn],
          BS_POS=
            akdes_tag_2024[infected_barn]),
      units=TRUE,
     plot = TRUE,
     verbose = TRUE,
     level.UD = 0.50,
     level = 0.95,debias = F,
     main="homerange 50% AKDE isopleth")

healthy_barn <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "RS" & blood_infection == "n")%>%
  pull(mean_AREA)

infected_barn <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "RS" & blood_infection == "y")%>%
  pull(mean_AREA)


# confidence levels

M1<-mean(healthy_barn)
M2 <-mean(infected_barn)


ratio<-M1/M2

S1<-sd(index_RS)
S2<-sd(index_MS)

N1<-length(index_RS)
N2<-length(index_MS)

# Calculate the Standard Error of the Log Ratio
SE_log_ratio <- sqrt((S1^2 / (N1 * M1^2)) + (S2^2 / (N2 * M2^2)))

# Output the Standard Error
lower_bound <- exp(log(ratio) - 1.96 * SE_log_ratio)
upper_bound <- exp(log(ratio) + 1.96 * SE_log_ratio)

# Output the confidence interval
c(lower_bound, upper_bound)

```

## 7.2 Meta Analysis House martin & infection

Here, we perform a meta analysis based on chi-square statistics (see Fleming et al. 2022)


```{r House martin & infection, cache=TRUE}
healthy_house <- AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "MS" & blood_infection == "n")%>%
  pull(index)

infected_house <- AKDE_size %>%
  mutate(index = row_number()) %>%
  filter(species == "MS" & blood_infection == "y") %>%
  pull(index)

# meta analysis
(meta(list(HM_NEG=
            akdes_tag_2024[healthy_house],
          HM_POS=
            akdes_tag_2024[infected_house]),
     plot = TRUE,
     verbose = TRUE,
     level.UD = 0.95,
     level = 0.95,debias = F,
     main="homerange 95% AKDE isopleth")->HM_meta_export_tag)

# akde 50%
meta(list(HM_NEG=
            akdes_tag_2024[healthy_house],
          HM_POS=
            akdes_tag_2024[infected_house]),
     plot = TRUE,
     verbose = TRUE,
     level.UD = 0.5,
     level = 0.95,debias = F,
     main="homerange 50% AKDE isopleth")


healthy_house <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "MS" & blood_infection == "n")%>%
  pull(mean_AREA)

infected_house <-  AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "MS" & blood_infection == "y")%>%
  pull(mean_AREA)


# confidence levels

M1<-mean(healthy_house)
M2 <-mean(infected_house)

S1<-sd(healthy_house)
S2<-sd(infected_house)

N1<-length(healthy_house)
N2<-length(infected_house)

ratio<-M1/M2


# Calculate the Standard Error of the Log Ratio
SE_log_ratio <- sqrt((S1^2 / (N1 * M1^2)) + (S2^2 / (N2 * M2^2)))

# Output the Standard Error
lower_bound <- exp(log(ratio) - 1.96 * SE_log_ratio)
upper_bound <- exp(log(ratio) + 1.96 * SE_log_ratio)

# Output the confidence interval
c(lower_bound, upper_bound)


```

# 7.3 Meta Analysis (all individuals) infected vs non-infected

```{r}
index_healthy <- AKDE_size %>%
    mutate(index = row_number())%>%
  filter(blood_infection == "n")%>%
  pull(index)

index_infected <- AKDE_size %>%
    mutate(index = row_number())%>%
  filter(blood_infection == "y")%>%
  pull(index)


healthy<-akdes_tag_2024%>%
  keep(seq_along(.) %in% index_healthy)

infected<-akdes_tag_2024%>%
  keep(seq_along(.) %in% index_infected)


(meta(list(non_infected=
            akdes_tag_2024[index_healthy],
          infected=
            akdes_tag_2024[index_infected]),
     plot = TRUE,
     IC="AICc",
     verbose = TRUE,
     level.UD = 0.95,
     level = 0.95,debias = TRUE,
     main="homerange 95% AKDE isopleth")->HM_meta_export_tag)
```

# Meta analysis swallow species

```{r}
index_RS <- AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "RS")%>%
  pull(index)

index_MS<- AKDE_size %>%
    mutate(index = row_number())%>%
  filter(species == "MS")%>%
  pull(index)


RS<-akdes_tag_2024%>%
  keep(seq_along(.) %in% index_RS)

MS<-akdes_tag_2024%>%
  keep(seq_along(.) %in% index_MS)



meta(list(RS=
            akdes_tag_2024[index_RS],
          MS=
            akdes_tag_2024[index_MS]),
     plot = TRUE,
     IC="AICc",
     verbose = TRUE,
     level.UD = 0.95,
     level = 0.95,debias = F,
     main="homerange 95% AKDE isopleth")

meta(list(RS=
            akdes_tag_2024[index_RS],
          MS=
            akdes_tag_2024[index_MS]),
     plot = F,
     IC="AICc",
     verbose = TRUE,
     level.UD = 0.5,
     level = 0.95,debias = TRUE,
     main="homerange 50% AKDE isopleth")

index_RS <- AKDE_size %>%
  rowwise() %>%
  filter(species == "RS") %>%
  mutate(index = row_number()) %>%
  pull(mean_AREA)

index_MS<- AKDE_size %>%
  rowwise() %>%
  filter(species == "MS") %>%
  mutate(index = row_number()) %>%
  pull(mean_AREA)


# confidence levels

M1<-mean(index_RS)
M2 <-mean(index_MS)

ratio<-M1/M2

S1<-sd(index_RS)
S2<-sd(index_MS)

N1<-length(index_RS)
N2<-length(index_MS)

# Calculate the Standard Error of the Log Ratio
SE_log_ratio <- sqrt((S1^2 / (N1 * M1^2)) + (S2^2 / (N2 * M2^2)))
exp(SE_log_ratio)


# Output the Standard Error
exp(log(ratio)-1.96*SE_log_ratio)
exp(log(ratio)+1.96*SE_log_ratio)


```

#Funnel plot to test if akde size is independent of 
```{r}
funnel(akdes_tag_2024,swallows_tag)
```

## 7 Simulation based on continuous time movement model and coordination manipulation algorithm. Short description:
ATLAS data is prone to data inconsistency, i.e. missing data occurs relatively often, and when it does, this is often related to the environment but not the tracking system itself (a common problem in line-of-sight tracking technologies). Here, we try to incorporate both, simulated movement data when animals are moving (characterized by shorter temporal gaps) and utilize the ctmm simulations for these small data gaps. In case of longer data gaps, we can be almost certain that the animal is not moving, or moving very little as not a single detection is coming through. In the latter cases, using a ctmm would estimate movement behaviour when in relality, there is not any movement at all. In the swallows, this is usually the case at their nesting colony, when individuals are resting / nesting at the colony, or sit on branches of trees to rest.

1.) we first save all true locations
2.) we simulate using the ctmm model through the data.

```{r}
# save true location
true_locations<-swallows_tag%>% 
  map_df(~ enframe(.x$t, name = NULL, value = "t"), .id = "tag_id")%>%
  mutate(location="real")

true<-split(true_locations$t,true_locations$tag_id)


# create some empty list
simulation_list<-list()
# simulation function
simulate_function <- function(i){
  
    # define sequence we need for prediction, fist observation of the day - last observation
  
    SEQ <- seq(from=true[[i]][1],
             to=true[[i]][length(true[[i]])],
             by=8)
  
   # predict missing locations using the ctmm model we fitted above (per individual, per day) 
  SIM <- simulate(swallows_tag[[i]],
                  ctmm_fit_tag[[i]],
                  t=SEQ)
  
  return(SIM)
  
}
### Parallelize 

cl <- parallel::makeCluster(30)
doParallel::registerDoParallel(cl)

simulation_list<- foreach(i=1: length(swallows_tag),.packages='ctmm',.errorhandling = 'remove') %dopar% { simulate_function(i) }



names(simulation_list)<-names(swallows_tag)


datalist<-list()

# data frame
for (i in 1:length(simulation_list)){
  
  t <-simulation_list[[i]]$t
  x <-simulation_list[[i]]$x
  y <-simulation_list[[i]]$y
  
  tag_id<-names(simulation_list[i])
  
  dataframe <- data.frame(tag_id,t,x,y)
  
  datalist[[i]] <- dataframe

}

simulation<- do.call(rbind, datalist)%>%
  left_join(true_locations,by=c("tag_id","t"))


```

3.) Biological filtering: we - again - remove all observations during night and omit the first 6 hours of tagging (capture induced biases)

```{r}
# spatial
swallows_ctmm_sim <- simulation%>%
  mutate(TIME=as.POSIXct(t,origin = "1970-01-01",tz="UTC"))%>%
  relocate(tag_id)%>%
  st_as_sf(coords = c("x","y"), crs = 3035,remove = FALSE)%>%
  mutate(long = st_coordinates(st_transform(.,4326))[,1],
         lat = st_coordinates(st_transform(.,4326))[,2])


sunlight_time<-suncalc::getSunlightTimes(
  date=lubridate::date(swallows_ctmm_sim$TIME),
                          lat=mean(swallows_ctmm_sim$lat),
                          lon=mean(swallows_ctmm_sim$long),
                        keep=c("dawn","dusk"),tz="GMT")%>%
  mutate(dawn=ymd_hms(dawn))%>%
  mutate(dusk=ymd_hms(dusk))%>%
  mutate(interval=interval(dawn,dusk))

swallows_tag_day<-
  swallows_ctmm_sim%>%
  tibble::add_column(sunlight_time$interval)%>%
  #keep just daylight observations
  filter(TIME %within% sunlight_time$interval)%>%
  # drop geometry columns
  st_drop_geometry()%>%
  dplyr::select(-c("sunlight_time$interval"))


s<-swallows_tag_day%>%
  group_by(tag_id)%>%
    mutate(tagtime_sum=max(TIME)-min(TIME),
         tagtime=as.numeric(TIME-min(TIME)))%>%
  ungroup()%>%
  filter(tagtime>(6*60*60))%>%
  ungroup()%>%
  mutate(location = replace_na(location, "filled"))

```
4.) we create the coordination manipulation function for larger data gaps
5.) applying the function to the data set (see code below) whenever there are subsequent 25 observations missing (i.e. 25*8 seconds = 224 secs = 3.73 min)
```{r}
coord_manipulation_function<-function(a,m){
  
  new_coord<- runif(min=(a-m/2),max=(a+m/2), n=1)

  return(new_coord)
}


df <- s %>%
  # Group by tag_id to perform operations within each unique tag_id
  group_by(tag_id) %>%
  # Add a column 'row_num' that indicates the row number within each tag_id group
  mutate(row_num = row_number()) %>%
  # Identify the row number of the first 'real' location within each tag_id group
  # NA_integer_ is used to handle non-'real' locations, na.rm = TRUE to ignore these NAs
  mutate(first_real_row = min(ifelse(location == "real", row_num, NA_integer_), na.rm = TRUE)) %>%
  # Filter out rows that are before the first 'real' location in each tag_id group
  filter(row_num >= first_real_row) %>%
  # Remove the helper columns 'row_num' and 'first_real_row' as they are no longer needed
  select(-row_num, -first_real_row) %>%
  # Regroup by tag_id for subsequent operations
  group_by(tag_id) %>%
  # Create a 'group' identifier for consecutive 'location' values using run-length encoding (rle)
  mutate(group = with(rle(location), rep(seq_along(lengths), lengths))) %>%
  # Group by both tag_id and the newly created 'group' to determine group lengths
  group_by(tag_id, group) %>%
  # Calculate the length of each group
  mutate(group_length = n()) %>%
  # Regroup by tag_id to carry forward the last 'real' x and y values
  group_by(tag_id) %>%
  # Assign the current 'real' x and y values, or NA if the location is not 'real'
  mutate(
    last_real_x = ifelse(location == "real", x, NA_real_),
    last_real_y = ifelse(location == "real", y, NA_real_)
  ) %>%
  # Handle the case where the first value is NA for 'last_real_x' and 'last_real_y'
  # If the first value is NA, replace it with the current x and y values
  mutate(
    last_real_x = ifelse(is.na(last_real_x) & row_number() == 1, x, last_real_x),
    last_real_y = ifelse(is.na(last_real_y) & row_number() == 1, y, last_real_y),
    # Carry forward the last known 'real' x and y values
    last_real_x = zoo::na.locf(last_real_x, na.rm = FALSE),
    last_real_y = zoo::na.locf(last_real_y, na.rm = FALSE)
  ) %>%
  # Un-group for final manipulation
  ungroup()%>%
  # Manipulate x and y values for 'filled' locations with a group_length >= 25
  # Using a custom function 'coord_manipulation_function'
  group_by(tag_id,t)%>%
mutate(
    x = if_else(location == "filled" & group_length >= 25, coord_manipulation_function(last_real_x, 10), x),
    y = if_else(location == "filled" & group_length >= 25, coord_manipulation_function(last_real_y, 10), y)
  )




saveRDS(df,
        here("output",
             "data-proc",
             "swallows_final_2024.RDS"))
```


# AKDE overlap calculations

```{r}
ctmm_overlap <- overlap(ctmm_fit_tag)
akde_overlap <- overlap(AKDES_tag_weights)

non_infected<-cmr%>%filter(blood_infection=="n")%>%pull(tag_id)
infected<-cmr%>%filter(blood_infection=="y")%>%pull(tag_id)


h_o<-overlap(AKDES_tag_weights%>%keep(names(.) %in% non_infected))
i_o<-overlap(AKDES_tag_weights%>%keep(names(.) %in% infected))

mean(h_o$CI[,,"est"],na.rm=T)
sd(h_o$CI[,,"est"],na.rm=T)

mean(i_o$CI[,,"est"],na.rm=T)
sd(i_o$CI[,,"est"],na.rm=T)


# infected individuals
RS_infected<-cmr%>%
  filter(blood_infection=="y" & species =="RS")%>%
  pull(tag_id)

RS_non_infected<-cmr%>%
  filter(blood_infection=="n" & species =="RS")%>%
  pull(tag_id)


overlap_RS_infected <- overlap(AKDES_tag_weights%>%keep(names(.) %in% RS_infected))

overlap_RS_non_infected <- overlap(AKDES_tag_weights%>%keep(names(.) %in% RS_non_infected))


mean(overlap_RS_infected$CI[,,"est"],na.rm=T)
mean(overlap_RS_infected$CI[,,"low"],na.rm=T)
mean(overlap_RS_infected$CI[,,"high"],na.rm=T)


mean(overlap_RS_non_infected$CI[,,"est"],na.rm=T)
mean(overlap_RS_non_infected$CI[,,"low"],na.rm=T)
mean(overlap_RS_non_infected$CI[,,"high"],na.rm=T)


# House martin
# infected individuals
MS_infected<-cmr%>%
  filter(blood_infection=="y" & species =="MS")%>%
  pull(tag_id)

MS_non_infected<-cmr%>%
  filter(blood_infection=="n" & species =="MS")%>%
  pull(tag_id)


overlap_MS_infected <- overlap(AKDES_tag_weights%>%keep(names(.) %in% MS_infected))

overlap_MS_non_infected <- overlap(AKDES_tag_weights%>%keep(names(.) %in% MS_non_infected))


mean(overlap_MS_infected$CI[,,"est"],na.rm=T)
mean(overlap_MS_infected$CI[,,"low"],na.rm=T)
mean(overlap_MS_infected$CI[,,"high"],na.rm=T)


mean(overlap_MS_non_infected$CI[,,"est"],na.rm=T)
mean(overlap_MS_non_infected$CI[,,"low"],na.rm=T)
mean(overlap_MS_non_infected$CI[,,"high"],na.rm=T)



########

m2<-akde_overlap$CI[,,"est"]

corrplot::corrplot(m2,order="FPC",type = "upper")


melted<-reshape2::melt(m2)%>%
  mutate(Var1=as.factor(Var1),
         Var2=as.factor(Var2)
         )


ggplot(data = melted, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile()+
  viridis::scale_fill_viridis(discrete=FALSE,
                              name= "Bhattacharyya coefficient")+
  d6::theme_d6()+
  labs(x=element_blank(),
       y=element_blank())+
  theme(axis.text.x = element_text(angle = 90),
        legend.key.width=unit(2,"cm"))


#ggsave(here("plots","overlap.png"), width = 20, height = 20, units = "cm",bg="white")
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
